---
title: "Nigeria LSMS All"
author: Elliot Quan
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(survey)
library(reshape2)
library(readr)
library(dplyr)
library(tidyr)
devtools::load_all() # load package functions
```


```{r}
# random functions: documentation is not 100% complete, will fix later.
# 
# auxiliary function for handling sums of NA vectors
# handles NA by returning NA if all are NA
sumNA <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm=TRUE))
  }
}

# take a vector of inputs: if at least 1 input is 1, then return 1 else 0.
# handles NA by returning NA if all are NA
indicator <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else if (sum(x, na.rm=TRUE) > 0) {
    return(1)
  } else {
    return(0)
  }
}

#' Converts a column of quantity data to kg based on unit coding provided
#' by units. 1 = kg, 2 = g, 3 = liter, 4 = centiliter. Does not mutate original df.
#'
#' @param quant quantity (character) 
#' @param units specifies units of quant (integer)
#' @param data (dataframe)
#'
#' @return converted quant vector
unit_conv <- function(quant, units) {
  return(case_when(
    units == 1 ~ quant,
    units == 2 ~ quant/1000,
    units == 3 ~ quant,
    units == 4 ~ quant/100
  ))
}

#' Same as above function, except coding 1 = liter, 2 = gram, 3 = kg, other = 0 (cannot convert units)
#' This also works for 1 = kg, 2 = gram, 3 = liter
unit_conv_pest <- function(quant, units) {
  return(case_when(
    units == 1 ~ quant,
    units == 2 ~ quant/1000,
    units == 3 ~ quant,
    is.numeric(units) ~ 0 
  ))
}

#' If x is 2 or 0, set to 0. This recodes an x column.
recode <- function(x) ifelse(x == 2 | x == 0, 0, 1)

#' this is mainly used for taking a y/n column x, and wanting to return 0 if x is 0 else y for each observation.
#' for example: if x = (0, 0, 1) and y = (NA, NA, 15), then we want to return (0, 0, 15).
handleNA <- function(x, y) ifelse(x == 0, 0, y)
```


# Description

This notebook generates a csv file which contains national level estimates of fertilizer use.

# Fertilizer data

For wave1/2, fert data is from the Post-Planting visit (so we need to use the weight associated with planting). For wave3, fert data is from the Post-Harvest visit (so we need to use the weight associated with harvest). 



```{r}
weights = load_csv_from_googledrive('1Ii83rJrX5dze6-wbNpCVj82sDHdHfXJY')
pw = weights %>% select(ea,hhid, w1 = wt_w1v1, w2 = wt_w2v1, w3 = wt_w3v2)
```



For some reason, some households that were sampled in wave2 planting are missing weights. We must then impute those weights (since their weight is equal to the weight of other households in the same enumeration area). This will also assign weights to wave2 households that were not interviewed, which is fine since those will be excluded. 10 household in w2 could not have their weights imputed, so they were dropped.

```{r}
pw2 = pw[, c('hhid', 'ea', 'w2')]
for (i in 1:nrow(pw2)) {
  if (is.na(pw2[i, 'w2'])) {
    # subset of pw2 with same ea
    same_ea = (pw2 %>% filter(ea == pw2$ea[i]))
    # find first non NA weight
    non_na_index = which(!is.na(same_ea$w2))[1]
    # impute the missing weight
    pw2[i, 'w2'] = same_ea[non_na_index, 'w2']
  }
}
pw$w2 = pw2$w2
pw = pw %>% select(-ea)
```

```{r}
pwtall = melt(pw, id.vars='hhid') %>% rename(start_yr = variable, weight = value)
vals = c(w1 = 2010, w2 = 2012, w3=2015)
pwtall$start_yr = dplyr::recode(pwtall$start_yr, !!!vals)
```



```{r}
#sect11d_platingw1.dta 
fert_w1 <- load_dta_from_googledrive('1P0I9YUBoRc8h93icSQ_f5Du6optQDIpW')
#sect11d_plantingw2.dta
fert_w2 <- load_dta_from_googledrive('1xCKxiEdTzpEaGFxZ9DgxjGGBitLq2WT-')
#secta11d_harvestw3.csv
fert_w3 <- load_csv_from_googledrive('1yNClVc5VQ7YB8ujxC1uZFSLU_WZEE1d2')
```

area data - need to join by plotid. area is in m^2, so need to convert to hectares (divide by 10000) later
```{r}
#sect11a1_plantingw1.dta
area_w1 <- load_dta_from_googledrive('1iG1CPzObtP9vBWJBzpEa1Mps3nMtAJqP')
#sect11a1_plantingw2.dta
area_w2 <- load_dta_from_googledrive('1v7LOdon8FiHXz6Hef_B6EZo9-_Ylw2Vj')
#sect11a1_plantingw3.csv
area_w3 <- load_csv_from_googledrive('1kUSspoxLRWS9a7sOp1-45cvwtI-pWhL0')

area_w1 <- area_w1 %>% select(hhid, plotid, area = s11aq4d) 
area_w2 <- area_w2 %>% select(hhid, plotid, area = s11aq4c)
area_w3 <- area_w3 %>% select(hhid, plotid, area = s11aq4c)

fert_w1 <- fert_w1 %>% inner_join(area_w1, by = c('hhid', 'plotid'))
fert_w2 <- fert_w2 %>% inner_join(area_w2, by = c('hhid', 'plotid'))
fert_w3 <- fert_w3 %>% inner_join(area_w3, by = c('hhid', 'plotid'))
fert_w3 = as.data.frame(fert_w3) 
```




Note: in wave3 organic fert is coded differently, but since we don't really care about it (and it is unclear what type is used) we will omit it. THere are 4 categories of fert: leftover (from previous planting seasons), free (from government/etc), bought1 (most purchased), and bought2 (second most purchased).

```{r}
fert1 <- fert_w1 %>% select(zone, state, hhid, plotid, area, fert_yn = s11dq1, 
                          leftover_yn = s11dq2, leftover_type = s11dq3, leftover_quant = s11dq4, 
                          free_yn = s11dq6, free_type = s11dq7, free_quant = s11dq8,
                          bought1_yn = s11dq12, bought1_type = s11dq14, bought1_quant = s11dq15,
                          bought2_yn = s11dq23, bought2_type = s11dq25, bought2_quant = s11dq26) %>% drop_na(fert_yn)


fert2 <- fert_w2 %>% select(zone, state, hhid, plotid, area, fert_yn = s11dq1, 
                            leftover_yn = s11dq2, leftover_type = s11dq3, leftover_quant = s11dq4,
                            free_yn = s11dq6, free_type = s11dq7, free_quant = s11dq8,
                            bought1_yn = s11dq12, bought1_type = s11dq15, bought1_quant = s11dq16,
                            bought2_yn = s11dq24, bought2_type = s11dq27, bought2_quant = s11dq28) %>% drop_na(fert_yn)

# wave3 quantities are varied, so need to standardize all weights to kg using unit_conv
fert3 <- fert_w3 %>% select(zone, state, hhid, plotid, area, fert_yn = s11dq1, 
                            leftover_yn = s11dq2, leftover_type = s11dq3, leftover_quant = s11dq4a,
                            leftover_quant_units = s11dq4b,
                            free_yn = sect11dq6, free_type = sect11dq7, free_quant = sect11dq8a,
                            free_quant_units = sect11dq8b,
                            bought1_yn = s11dq12, bought1_type = s11dq15, bought1_quant = s11dq16a,
                            bought1_quant_units = s11dq16b,
                            bought2_yn = s11dq24, bought2_type = s11dq27, bought2_quant = s11dq28a,
                            bought2_quant_units = s11dq28b) %>% drop_na(fert_yn) 

fert3 <- fert3 %>% mutate(leftover_quant = unit_conv(leftover_quant, leftover_quant_units),
                          free_quant = unit_conv(free_quant, free_quant_units),
                          bought1_quant = unit_conv(bought1_quant, bought1_quant_units), 
                          bought2_quant = unit_conv(bought2_quant, bought2_quant_units)) %>% select(-contains('units'))

fert1
fert2
fert3
```

```{r}
w1_fert <- data.frame(country="nigeria", start_yr = 2010, end_yr = 2011, fert1) 
w2_fert <- data.frame(country="nigeria", start_yr = 2012, end_yr = 2013, fert2)
w3_fert <- data.frame(country="nigeria", start_yr = 2015, end_yr = 2016, fert3)
fert = rbind(w1_fert, w2_fert, w3_fert)
```

```{r}
# convert m^2 to hectares
fert$area = fert$area/10000
```





# recoding data
```{r}
# for all y/n cols, change 2s to 0s (1 = yes, 0 = no)
fert <- fert %>% mutate_at(c('fert_yn', 'leftover_yn', 'free_yn', 'bought1_yn', 'bought2_yn'), recode)

# if no fert used, then set the following cols to 0 instead of NA. See handleNA fxn above.
fert <- fert %>% mutate_at(c('leftover_yn', 'free_yn', 'bought1_yn', 'bought2_yn', 'leftover_type', 'free_type', 'bought1_type', 'bought2_type', 'leftover_quant', 'free_quant', 'bought1_quant', 'bought2_quant'), list(~handleNA(fert_yn, .)))

# same goes for the other categories of use: if none used, set both type and quant to 0.
fert <- fert %>% mutate_at(c('leftover_type', 'leftover_quant'), list(~handleNA(leftover_yn, .))) %>% 
  mutate_at(c('free_type', 'free_quant'), list(~handleNA(free_yn, .))) %>%
  mutate_at(c('bought1_type', 'bought1_quant'), list(~handleNA(bought1_yn, .))) %>%
  mutate_at(c('bought2_type', 'bought2_quant'), list(~handleNA(bought2_yn, .)))
```


```{r}
# for each of leftover, free, bought1, bought2, check that the type of fertilizer matches the desired type. if so, then add that quantity to the sum and return the total quantities for the type.
agg_quant <- function(type, df) {
  return(ifelse(df$leftover_type == type, df$leftover_quant, 0) + ifelse(df$free_type == type, df$free_quant, 0) + ifelse(df$bought1_type == type, df$bought1_quant, 0) + ifelse(df$bought2_type == type, df$bought2_quant, 0))
}

# return 1 if at least one of the leftover, free, bought1, bought2 equals type, else 0.
yn_quant <- function(type, df) {
  return(as.numeric(((df$leftover_type == type) | (df$free_type == type) | (df$bought1_type == type) | (df$bought2_type == type))))
}

fert <- fert %>% mutate(npk_yn  = yn_quant(1, fert), urea_yn = yn_quant(2, fert), comp_manure_yn = yn_quant(3, fert), other_yn = yn_quant(4, fert))

fert <- fert %>% mutate(npk_quant = agg_quant(1, fert), urea_quant = agg_quant(2, fert), comp_manure_quant = agg_quant(3, fert), other_quant = agg_quant(4, fert)) 
```


```{r}
# add indicator vars for organic (ofert), inorganic (iofert), all fert (fert)
fert1 <- fert  %>% mutate(ofert_yn = comp_manure_yn, iofert_yn = as.numeric(select(., matches('npk_yn|urea_yn')) %>% rowSums(na.rm = TRUE) > 0)) 
fert1
```

```{r}
# final selection
fert_df <- fert1 %>% select(start_yr, hhid, area, fert_yn, npk_yn, urea_yn, comp_manure_yn, other_yn, ofert_yn, iofert_yn, npk_quant, urea_quant, comp_manure_quant, other_quant)

```

```{r}
fert_df$total_quant = fert_df %>% select(contains('quant')) %>% rowSums(na.rm=TRUE)
```

Now we group the data by household. Quantity & area (continuous) will be summarized by sum, indicator variables (categorical) will be summarized by an indicator.

```{r}
fert_cont = fert_df %>% select(start_yr, hhid, matches('area|quant')) %>% group_by(start_yr, hhid) %>% summarize_all(sum, na.rm=TRUE)
fert_cat = fert_df %>% select(-matches('area|quant')) %>% group_by(start_yr, hhid) %>% summarize_all(indicator)
fert_df = fert_cont %>% inner_join(fert_cat)
```




Now we aggregate the data for each of the 3 surveys and use the survey weights to produce national lvl estimates. Note that hhids are not unique across surveys (since households are revisited), so we need to join by hhid and start_yr.


```{r}
fert = fert_df %>% inner_join(pwtall, by=c('start_yr', 'hhid'))
```



```{r}
fw1 = fert %>% filter(start_yr == 2010)
fw2 = fert %>% filter(start_yr == 2012) %>% drop_na(weight)
fw3 = fert %>% filter(start_yr == 2015)
fw1des = svydesign(id = ~hhid,
                        weights = ~weight,
                        nest = TRUE,
                        data = fw1)
fw2des = svydesign(id = ~hhid,
                        weights = ~weight,
                        nest = TRUE,
                        data = fw2)
fw3des = svydesign(id = ~hhid,
                        weights = ~weight,
                        nest = TRUE,
                        data = fw3)
```


```{r}

cols = c('fert_yn', 'npk_yn', 'urea_yn', 'comp_manure_yn', 'other_yn', 'ofert_yn', 'iofert_yn', 'npk_quant', 'urea_quant', 'comp_manure_quant', 'other_quant', 'total_quant', 'area')
d = length(cols)

calc_res = function(design) {
  res = c(0, d)
  for (var in 1:d) {
    f = as.formula(paste('~', cols[var]))
    if (grepl('quant|area', cols[var])) {
      res[var] = svytotal(f, design, na.rm=TRUE)
    } else {
      res[var] = svymean(f, design, na.rm=TRUE)
    }
    
  }
  return(res)
}
res = matrix(0, nrow=3, ncol=d)
res[1,] = calc_res(fw1des)
res[2,] = calc_res(fw2des)
res[3,] = calc_res(fw3des)

colnames(res) = cols
rownames(res) = c(2011, 2013, 2015)
res
```




# Pesticide data (revisit later)

Note: Wave2 is missing sect11c2_plantingw2, so until that can be found we will omit for now.

```{r}
#sect11c_plantingw1.csv
w1_pest <- load_dta_from_googledrive('1thrw4617geattiMhnkpkgy-e4h19jG8O')
w1_pest

#secta11c2_harvestw3.csv
w3_pest <- load_csv_from_googledrive('1NA_7J10lTjSFS4K13egPtWMYqixcijuf')
w3_pest
```



```{r}
pest1 <- w1_pest %>% select(hhid, plotid, pest_yn = s11cq1, pest_quant = s11cq2a, pest_quant_unit = s11cq2b, herb_yn = s11cq10, herb_quant = s11cq11a, herb_quant_unit = s11cq11b) %>% drop_na(pest_yn, herb_yn)
pest3 <- w3_pest %>% select(hhid, plotid, pest_yn =s11c2q1, pest_quant = s11c2q2a, pest_quant_unit = s11c2q2b, herb_yn = s11c2q10, herb_quant = s11c2q11a, herb_quant_unit = s11c2q11b) %>% drop_na(pest_yn, herb_yn)
pest1
pest3
```


```{r}
w1_p <- data.frame(country="nigeria", start_yr = 2010, end_yr = 2011, pest1, stringsAsFactors = F) 
w3_p <- data.frame(country="nigeria", start_yr = 2015, end_yr = 2016, pest3, stringsAsFactors = F)
pest_c <- rbind(w1_p, w3_p)
pest_c
```



```{r}
# for y/n cols, set 2 to 0
pest_c[pest_c$pest_yn == 2, c('pest_yn')] <- 0
pest_c[pest_c$herb_yn == 2, c('herb_yn')] <- 0

# if no pest used or no herb used, then set quant and type columns to 0.
pest_c <- pest_c %>% mutate_at(c('pest_quant', 'pest_quant_unit'), list(~handleNA(pest_yn, .))) %>% 
  mutate_at(c('herb_quant', 'herb_quant_unit'), list(~handleNA(herb_yn, .)))
pest_c

pest_df <- pest_c %>% mutate(pest_quant = unit_conv_pest(pest_quant, pest_quant_unit), herb_quant = unit_conv_pest(herb_quant, herb_quant_unit))
pest_df
```

# Merging Fertilizer + Pesticide

merge on country, start_yr, end_yr, hhid, plotid. To preserve the wave2 fertilizer data, we do a left join (since w2 pesticide data is missing).

```{r}
final_df <- fert_df %>% left_join(pest_df, by = c('country', 'start_yr', 'end_yr', 'hhid', 'plotid'))
final_df 
```

```{r}
final <- final_df %>% select(country, start_yr, end_yr, zone, state, hhid, plotid,  fert_rep_yn, npk_yn, urea_yn, comp_manure_yn, other_yn, ofert_yn, iofert_yn, pest_yn,  herb_yn,  quant_npk = npk_quant, quant_urea = urea_quant, quant_comp_manure = comp_manure_quant, quant_other = other_quant,  quant_pest = pest_quant, quant_herb = herb_quant, area, longitude, latitude) %>% select(-c('area', 'longitude', 'latitude'), c('area', 'longitude', 'latitude'))
final
```

```{r, eval=FALSE}
# write the df only 
write_csv(final, '../results/nga_lsms_full.csv')
```






