---
title: "LSMS Uganda Full"
author: Elliot Quan
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(survey)
library(sf)
library(tidyr)
library(dplyr)
library(readr)
library(stringr)
library(dummies)
devtools::load_all()
```

```{r}
# auxiliary function for handling sums of NA vectors
# handles NA by returning NA if all are NA
sumNA <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm=TRUE))
  }
}

# take a vector of inputs: if at least 1 input is 1, then return 1 else 0.
# handles NA by returning NA if all are NA
indicator <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else if (sum(x, na.rm=TRUE) > 0) {
    return(1)
  } else {
    return(0)
  }
}

#' Converts a column of quantity data to kg based on unit coding provided
#' by units. 1 = kg, 2 = g, 3 = liter, 4 = centiliter. Does not mutate original df.
#'
#' @param quant quantity (character) 
#' @param units specifies units of quant (integer)
#' @param data (dataframe)
#'
#' @return converted quant vector
unit_conv <- function(quant, units) {
  return(case_when(
    units == 1 ~ quant,
    units == 2 ~ quant/1000,
    units == 3 ~ quant,
    units == 4 ~ quant/100
  ))
}

#' If x is 2 or 0, set to 0. This recodes an x column. Note that NAs are recoded to 0.
recode <- function(x) ifelse(is.na(x) | x == 2 | x == 0, 0, 1)

#' this is mainly used for taking a y/n column x, and wanting to return 0 if x is 0 else y for each observation.
#' for example: if x = (0, 0, 1) and y = (NA, NA, 15), then we want to return (0, 0, 15).
handleNA <- function(x, y) ifelse(x == 0, 0, y)
```



# Description

This notebook generates plot level fertilizer information.
To keep consistency, we will only analyze B files. A files denote the first visit (Jan-Jun), while B files denote the second visit (Jul-Dec) in the year. This is because only 1/2 of the A files were intact, so we'll just use the B files (since all are intact).


We will also analyze 09 last since its format differs slightly.




Plot Area data is in AGSEC4, so we'll need those too.  
```{r}
#AGSEC3B
w1_3B <- load_dta_from_googledrive('19Ah0MdOilhUqd-lrLXIbeklcvC-lTZGU')
w2_3B <- load_dta_from_googledrive('1BVQD632HvuyVSUuGMojwWa-49OvZmscx')
w3_3B <- load_dta_from_googledrive('1H4xv_8xMMS1E794ILtAFNHCIngbqy89P')
w4_3B <- load_csv_from_googledrive('1d0y5MZtQmR6FTiS_nu5ETdVgCSV4rI3l')

#AGSEC4B
w1_4B <- load_dta_from_googledrive('1A73gVw8ZW2jQoktHowKtpMnOtfm5yYP9')
w2_4B <- load_dta_from_googledrive('1J77rX_YqiUrfcZwFka5G9DlwnUd84DXs')
w3_4B <- load_dta_from_googledrive('1-PJACIKbzvaMHzsY0Yo8bQ0ackBtNCuH')
w4_4B <- load_csv_from_googledrive('1TSF51Buqy6Vtv_zZkzLMzkPz6wdI4RqE')

#UNPS_GEOVARS_xxxx.dta
w1_geo <- load_dta_from_googledrive('1DfdCow-OWianLigs9tnaVGQY8np73fK_')
w2_geo <- load_dta_from_googledrive('1zy5I8KIxrN5v1FM2fwverG4mvjzKy5-F')
w3_geo <- load_dta_from_googledrive('1kGQdP0NTGp4I--gjcBZmzDJf4yTvZpOS')

# Weight data: GSEC1.dta
w1_w = load_dta_from_googledrive('1653pmufjZDtxTZ3uCohb74rkUB_eggMl')
w2_w = load_dta_from_googledrive('1-YE9mG85w5-VSgsEIDx9ihwLcSVtS3vm')
w3_w = load_dta_from_googledrive('1ikEk0ejSIz-_8YDhdaXhIfPtIELkKQ_3')
w4_w = load_csv_from_googledrive('1F7vwOz_OZKzb6IkgLZBrI_cdRKI1ZL1u')
```

```{r}
w1_weight = w1_w %>% select(HHID, weight=wgt09) %>% mutate(start_yr=2009)# wgt09
w2_weight = w2_w %>% select(HHID, weight=wgt10) %>% mutate(start_yr = 2010)# wgt10
w3_weight = w3_w %>% select(HHID, weight= mult) %>% mutate(start_yr = 2011) # mult
w4_weight = w4_w %>% select(HHID, weight= wgt_X) %>% mutate(start_yr = 2015) # wgt_X
w4_weight$HHID = as.numeric(gsub("^0+", x=gsub('\\D+', x=w4_weight$HHID, ""), "")) # convert HHIDs back to numeric form

ugaweights = rbind(w1_weight, w2_weight, w3_weight, w4_weight)
```


# Fertilizer/Pesticide data

This includes reported org/inorg fert, quantities, and types if applicable. Quantities are coded as kg/liters, but I am making the assumption that 1 kg = 1 liter. 

```{r}
w1_fert <- w1_3B %>% select(hhid=HHID, parcelid=a3bq1, plotid=a3bq3,  ofert_rep_yn = a3bq4, quant_ofert = a3bq5, iofert_rep_yn = a3bq14, type_iofert = a3bq15, quant_iofert = a3bq16, pest_yn = a3bq26, quant_pest = a3bq28b) 

w2_fert <- w2_3B  %>% select(hhid=HHID, parcelid=prcid, plotid=pltid, ofert_rep_yn = a3bq4, quant_ofert = a3bq5, iofert_rep_yn = a3bq14, type_iofert = a3bq15, quant_iofert = a3bq16, pest_yn = a3bq26, quant_pest = a3bq28b)


w3_fert <- w3_3B %>% select(hhid=HHID, parcelid = parcelID, plotid=plotID, ofert_rep_yn = a3bq4, quant_ofert = a3bq5, iofert_rep_yn = a3bq13, type_iofert = a3bq14, quant_iofert = a3bq15, pest_yn = a3bq22, quant_pest = a3bq24b)

w4_fert <- w4_3B %>% select(hhid=HHID, parcelid=parcelID, plotid=plotID, ofert_rep_yn = a3bq4, quant_ofert = a3bq5, iofert_rep_yn = a3bq13, type_iofert = a3bq14, quant_iofert = a3bq15, pest_yn = a3bq22, quant_pest = a3bq24b)

```

Area data is one level deeper than the plot data, so each plot's area will be the sum of the areas of each 'subplot', where a subplot corresponds to a crop. Thus we need to group by plot_id. Also area is in acres.

```{r}
w1_area <- w1_4B %>% select(hhid=HHID, parcelid = a4bq2, plotid=a4bq4, area=a4bq8) %>% group_by(hhid, parcelid, plotid) %>% summarize(area=sum(area))

w2_area <- w2_4B %>% select(hhid=HHID, parcelid = prcid, plotid=pltid, area=a4bq8) %>% group_by(hhid, parcelid, plotid) %>% summarize(area=sum(area))

w3_area <- w3_4B %>% select(hhid=HHID, parcelid = parcelID, plotid = plotID, area=a4bq7) %>% group_by(hhid, parcelid, plotid) %>% summarize(area=sum(area))

w4_area <- w4_4B %>% select(hhid=HHID, parcelid=parcelID, plotid=plotID, area=a4bq7) %>% group_by(hhid, parcelid, plotid) %>% summarize(area=sum(area))
```


```{r}
w1 <- w1_fert %>% left_join(w1_area, by = c('hhid', 'parcelid', 'plotid'))
w2 <- w2_fert %>% left_join(w2_area, by = c('hhid', 'parcelid', 'plotid'))
w3 <- w3_fert %>% left_join(w3_area, by = c('hhid', 'parcelid', 'plotid'))
w4 <- w4_fert %>% left_join(w4_area, by = c('hhid', 'parcelid', 'plotid'))
```


```{r}
w1t <- data.frame(country='uganda', start_yr=2009, end_yr=2010, w1, stringsAsFactors = FALSE)
w2t <- data.frame(country='uganda', start_yr=2010, end_yr=2011, w2, stringsAsFactors = FALSE)
w3t <- data.frame(country='uganda', start_yr=2011, end_yr=2012, w3, stringsAsFactors = FALSE)
w4t <- data.frame(country='uganda', start_yr=2015, end_yr=2016, w4, stringsAsFactors = FALSE)
uga <- rbind(w1t, w2t, w3t, w4t)
```


# Transforming Fertilizer data

1) need to create dummy vars for each type of inorganic fert 
2) need to handle NAs for nonreported data

```{r}
uga
```


# Handling NAs

Since there is no question of 'was fertilizer used' to discern whether NAs in iofert/ofert correspond to 0s, we will make the assumption that NAs mean 0 here. This assumption was NOT made in the other LSMS surveys. This occurs in approximately 1.9k cases.

```{r}
uga[is.na(uga$ofert_rep_yn) | is.na(uga$iofert_rep_yn | is.na(uga$pest_yn)), c('ofert_rep_yn', 'iofert_rep_yn', 'pest_yn')]
```




```{r}
# we will recode all 2s to 1s. also, if no org. fert, inorg fert. or pest is used, then the corresponding quantities/types will be set to 0.
uga <- uga %>% mutate_at(c('ofert_rep_yn', 'iofert_rep_yn', 'pest_yn'), recode)

uga <- uga %>% mutate_at('quant_ofert', list(~handleNA(ofert_rep_yn, .))) %>% mutate_at(c('quant_iofert', 'type_iofert'), list(~handleNA(iofert_rep_yn, .))) %>% mutate_at('quant_pest', list(~handleNA(pest_yn, .)))

# handle more NAs (if fert was reported but the type is unreported, then categorize it as other/mixed (4))
uga[(uga$iofert_rep_yn == 1) & (is.na(uga$type_iofert)), ] <- 4
```


```{r}
# this function creates a quantity col for a corresponding fertilizer type.
agg_quant <- function(type, df) {
  return(ifelse(df$type_iofert == type, df$quant_iofert, 0))
}

# this function creates a type col for a corresponding fertilizer type
agg_yn <- function(type, df) {
  return(as.numeric((!is.na(df$type_iofert)) & (df$type_iofert == type)))
}

uga_mod <- uga %>% mutate(nitrate_quant=agg_quant(1, uga), phosphate_quant=agg_quant(2, uga), potash_quant=agg_quant(3, uga), mixed_quant=agg_quant(4, uga))

uga_mod <- uga_mod %>% mutate(nitrate_yn=agg_yn(1, uga), phosphate_yn=agg_yn(2, uga), potash_yn=agg_yn(3, uga), mixed_yn=agg_yn(4, uga))
```

Scale area (acres) to meters^2. Conv factor is 1 acre = 4046.86 m^2
```{r}
uga_mod$area <- uga_mod$area*4046.86
uga_final <- uga_mod
```

```{r}
uga<- uga_final %>% select(country, start_yr, end_yr, hhid, plotid, parcelid, ofert_rep_yn, quant_ofert, iofert_rep_yn, type_iofert, quant_iofert, pest_yn, quant_pest, nitrate_quant, phosphate_quant, potash_quant, mixed_quant, nitrate_yn, phosphate_yn, potash_yn, mixed_yn, area) %>% rename(quant_nitrate = nitrate_quant, quant_phosphate = phosphate_quant, quant_potash = potash_quant, quant_mixed = mixed_quant)
uga = uga %>% mutate(fert_dummy = as.numeric(iofert_rep_yn + ofert_rep_yn > 0))
uga
```

# Survey Weighting


```{r}
uga_joined = uga %>% inner_join(ugaweights, by = c('start_yr', 'hhid'='HHID'))
uga_joined = uga_joined %>% drop_na(weight) # 15 missing weights - we can try to impute later?
```

```{r}
sum(is.na(uga_joined$weight))
```

```{r}

w1 = uga_joined %>% filter(start_yr==2009)
w2 = uga_joined %>% filter(start_yr==2010)
w3 = uga_joined %>% filter(start_yr==2011)
w4 = uga_joined %>% filter(start_yr==2015)

pw1des = svydesign(id = ~hhid,
                        weights = ~weight,
                        nest = TRUE,
                        data = w1)
pw2des = svydesign(id = ~hhid,
                        weights = ~weight,
                        nest = TRUE,
                        data = w2)
pw3des = svydesign(id = ~hhid,
                        weights = ~weight,
                        nest = TRUE,
                        data = w3)
pw4des = svydesign(id= ~hhid,
                   weights = ~weight,
                   nest = TRUE,
                   data = w4)
```

```{r}
uga_joined
```


```{r}
cols = c('fert_dummy', 'ofert_rep_yn', 'iofert_rep_yn', 'pest_yn', 'quant_ofert', 'quant_iofert', 'quant_pest')  
d = length(cols)

calc_means = function(design) {
  means = c(0, d)
  for (var in 1:d) {
    f = as.formula(paste('~', cols[var]))
    means[var] = svymean(f, design, na.rm=TRUE)
  }
  return(means)
}
res1 = matrix(0, nrow=4, ncol=d)
res1[1,] = calc_means(pw1des)
res1[2,] = calc_means(pw2des)
res1[3,] = calc_means(pw3des)
res1[4,] = calc_means(pw4des)

colnames(res1) = cols
rownames(res1) = c(2009, 2010, 2011, 2015)
res1 = data.frame(res1)
```


```{r, eval=FALSE}
write_csv(res1, 'natl_agg_results/uga_res.csv')
```




